A Hierarchical Bayesian Framework for Inferring the Stellar Obliquity Distribution

### 2023-04-13 src/tex/ms.tex
[showyourwork] overleaf sync

To investigate the performance of our hierarchical Bayesian framework, we apply it to simulated data, generated with a known ground-truth stellar obliquity distribution.
\edits{We test the following four $\cos{\psi}$ distributions: a uniform distribution such that $\cos{\psi} \sim \mathcal{U}(-1,1)$, and three truncated Normal distributions bounded between -1 and 1, such that $\cos{\psi} \sim \mathcal{N}(0,0.2)$, $\cos{\psi} \sim \mathcal{N}(-0.4,0.2)$, and $\cos{\psi} \sim \mathcal{N}(0.4,0.2)$.}
For each $\cos{\psi}$ distribution, we randomly generate 200 samples of sky-projected stellar obliquity $\lambda$ and stellar inclination $i_\star$. We assume the stellar spin axis is uniformly distributed around the planetary orbital axis in the azimuthal direction and the orbital inclination is $90\degr$.
The sampled $i_\star$ and $\lambda$ here are \emph{true} values. 
To simulate the observation process, we add Gaussian noise to the \emph{true} $\lambda$ and $i_\star$, using uncertainties of $\sigma_{\lambda} = 8\degr$ and $\sigma_{i_\star} = 10\degr$, which are typical of the literature sample \cite{Albrecht22}.
Using these simulated $\lambda$ measurements and their uncertainties, we infer the $\cos{\psi}$ distribution of the sample with \edits{both $\lambda$ and $i_\star$ likelihoods, $\lambda$ likelihood only, and $i_\star$ likelihood only.}

@@

\begin{figure*}[ht!]
    \centering
    \script{sim_plot.py}
    \includegraphics[width=0.8\textwidth]{figures/simulation.pdf}
    \caption{Inferred stellar obliquity distributions from sky-projected stellar obliquities with and without information on the stellar inclination, depicted in blue and orange curves, respectively. Each row presents a set of simulated data with the true distribution of $\cos{\psi}$ indicated by grey dashed curves. The shallow contours represent the 1-$\sigma$ and 2-$\sigma$ uncertainties of the inferred distributions.}
    \label{fig:simulation}
\end{figure*}

@@

\begin{figure*}[ht!]
    \centering
    \script{transform.py}
    \includegraphics[width=0.8\linewidth]{figures/transform.pdf}
    \caption{Simulated $\cos{\psi}$ distributions (\nth{1} column) and the corresponding distributions of sky-projected stellar obliquity $\lambda$ (\nth{2} column) and stellar inclination $i_\star$ (\nth{3} column). The inferred $\cos{\psi}$ distributions assuming isotropic stellar inclinations are shown in the \nth{4} column. The grey histograms present the random samplings of $\lambda$ and $i_\star$ from the $\cos{\psi}$ distributions, and the blue curves present the numerical solutions.}
    \label{fig:transform}
\end{figure*}

>>>>

To investigate the performance of our hierarchical Bayesian framework, we apply it to simulated data, generated with a known ground-truth stellar obliquity distribution.
\edits{We test the following four $\cos{\psi}$ distributions: a uniform distribution bounded between $-1$ and $1$ such that $\cos{\psi} \sim \mathcal{U}(-1,1)$, and three truncated Normal distributions bounded between $-1$ and $1$, such that $\cos{\psi} \sim \textit{truncnorm}(0, 0.2)$, $\cos{\psi} \sim \textit{truncnorm}(-0.4, 0.2)$, and $\cos{\psi} \sim \textit{truncnorm}(0.4, 0.2)$, where the first and second number are the mean and standard deviation of the normal distribution, respectively.}
For each $\cos{\psi}$ distribution, we randomly generate 200 samples of sky-projected stellar obliquity $\lambda$ and stellar inclination $i_\star$. We assume the stellar spin axis is uniformly distributed around the planetary orbital axis in the azimuthal direction and the orbital inclination is $90\degr$.
The sampled $i_\star$ and $\lambda$ here are \emph{true} values. 
To simulate the observation process, we add Gaussian noise to the \emph{true} $\lambda$ and $i_\star$, using uncertainties of $\sigma_{\lambda} = 8\degr$ and $\sigma_{i_\star} = 10\degr$, which are typical of the literature sample \cite{Albrecht22}.
Using these simulated $\lambda$ measurements and their uncertainties, we infer the $\cos{\psi}$ distribution of the sample with both $\lambda$ and $i_\star$ likelihoods or $\lambda$ likelihood only.

@@

\begin{figure}[ht!]
    \centering
    \script{sim_plot.py}
    \includegraphics{figures/simulation.pdf}
    \caption{Inferred stellar obliquity distributions from sky-projected stellar obliquities with and without information on the stellar inclination, depicted in blue and orange curves, respectively. Each row presents a set of simulated data with the true distribution of $\cos{\psi}$ indicated by grey dashed curves. The shallow contours represent the 1-$\sigma$ and 2-$\sigma$ uncertainties of the inferred distributions.}
    \label{fig:simulation}
\end{figure}

@@

\begin{figure*}[ht!]
    \centering
    \script{transform.py}
    \includegraphics[width=0.7\linewidth]{figures/transform.pdf}
    \caption{Simulated $\cos{\psi}$ distributions (\nth{1} column) and the corresponding distributions of sky-projected stellar obliquity $\lambda$ (\nth{2} column) and stellar inclination $i_\star$ (\nth{3} column). The inferred $\cos{\psi}$ distributions assuming isotropic stellar inclinations are shown in the \nth{4} column. The grey histograms present the random samplings of $\lambda$ and $i_\star$ from the $\cos{\psi}$ distributions, and the blue curves present the numerical solutions.}
    \label{fig:transform}
\end{figure*}
